SoK: Defending Against Web Application Attacks: From Research Approaches to Practical Tools

### 2014-11-14 docs/questioning.tex
stylistic fixes

There is already a survey on mitigating software vulnerabilities
in general~\cite{SZ12}. The scope of that
research is much broader and leaves out the majority of the
approaches and mechanisms that we consider in this work.
This is mainly because that work includes countermeasures
that identify vulnerabilities using program analysis
(which takes place during the development or testing phases), while
we consider protection mechanisms that counter
web attacks when they take place. Similarly, static or
dynamic analysis techniques that examine applications to
identify vulnerabilities that may lead to the attacks
that we described earlier are out of scope as well.
While countermeasures against {\sc sql}
injection attacks have been previously surveyed in a 2006 paper~\cite{HVO06},
a significant number of new {\sc sql} injection defenses has been proposed since
then, as we mentioned earlier.

@@

The main vulnerability behind the most dangerous
web application attacks involves
the absence of input validation. By taking advantage of
this, attackers can inject their code into
applications to perform malicious tasks. Exploits
of this kind can have different forms depending on the
execution context of the application and the location
of the programming flaw that leads to the attack.

@@

Code injection attacks can be divided into two categories. The first
involves binary code and the second higher-level language code. An
extensive survey on binary code injection attacks was conducted by
Lhee and Chapin~\cite{LC03}. Advances in memory corruption vulnerability
exploitation have been studied extensively~\cite{PB04, SPWS13}
and countermeasures to such attacks
have already been surveyed~\cite{YJP12}\cite[Sec. 13.8]{DKZ12}.
In this work
we do not consider binary code injection, focusing instead on defenses
that protect web applications against attacks based on the
injection of higher level language code.

Figure~\ref{fig:taxonomy} presents a taxonomy of source
code injection attacks against web applications.
Such attacks may involve high-level language code, written in either a
{\it Domain Specific Language} ({\sc dsl}) or a {\it Dynamic Language}.
In the following, we briefly discuss examples from both categories
that will be used throughout the paper.

Injection attacks that involve {\sc dsl}s constitute an important
subset of the code injection problem, as {\sc dsl}s such as {\sc sql} and
{\sc xml} play an significant role in the development of both web and
mobile applications. For example, many applications have interfaces
through which a user enters input to interact with the application,
thereby interacting with the underlying database. This input can
become part of an {\sc sql} query and get executed on the target
database. Code injection attacks that exploit vulnerabilities in database
interfaces by taking advantage of input validation issues, such as
incorrectly passed parameters or incorrect type handling, are called
{\sc sql} injection attacks~\cite{CERT02,MS09,HVO06,SW06}.
Consider a trivial example that takes advantage of incorrectly
filtered quotation characters in an application that
sends a temporary password for a forgetful user via email by executing the
following query:

@@

\noindent
Attackers that would input the string {\tt anything' OR 'x'='x},
could obtain the administrator's password by email.
Savvy programmers can use certain {\sc api} functions, such as {\sc php}'s
{\tt mysql\_real\_escape\_string()} to detect malformed input, or
better use prepared {\sc sql} statements, instead of statement
templates. Unfortunately,
the increasing number of {\sc sql} injection attacks
suggest that programmers are not always that careful.
Using similar techniques, malicious users can mount other exploits based on
{\sc dsl}s such as {\sc xp}ath~\cite{SW06,CDL07,MKS09},
{\sc xml}~\cite{MSM13} and {\sc json}~\cite{SMS13}. With this kind of
attacks, a malicious user can view sensitive information, destroy or
modify protected data, or even crash the entire application.

{\sc html} is another {\sc dsl} that can be used for malicious purposes
when an application does not properly handle user-supplied data. Based
on this vulnerability attackers can supply valid {\sc html},
typically via a parameter value, and inject their own content into the
page. {\sc html} injection is mainly associated with {\sc xss}
attacks~\cite{BJM08,LZRL09}. However, it can also be used as a
vehicle for cross-site request forgery attacks in the following manner:
Consider a vulnerable bulletin board system where {\tt img} tags are not checked.
A malicious user could embed a {\sc csrf} request within an {\tt img}
tag in the following manner:

@@

JavaScript injection attacks make up a large subset of dynamic
language code injection attacks. Such attacks are enabled when
a web application accepts
and redisplays data of uncertain origin without
proper validation and filtering. Based on this flaw, an attacker
can manage to inject a script in the JavaScript engine of a browser
and alter its execution flow~\cite{ELX07}.
JavaScript injection attacks are considered a critical issue
in web application security mainly because it is associated with
major vulnerabilities such as {\sc xss} attacks~\cite{SG07} and
Cross-Channel Scripting ({\sc xcs}) attacks~\cite{W10,BBB09}.
A typical {\sc xss} example that involves JavaScript injection
is the following: consider a web page that prints the value
of a query parameter ({\tt query}) from the
page's {\sc url} as part of the page's content
without escaping the value. Attackers
can take advantage of this and inject an {\tt iframe} tag
into the page to steal a user's cookie and
send it via an image request to a web site
under their control (malicious.com).
This could be achieved by including the following
link to the malicious web site (or send it via phishing
email) and induce the user to click on it:

@@

In this section we provide a step-by-step exploitation model to
understand the process of carrying out source code injection attacks
against web applications.
Figure~\ref{fig:attacks}
presents the required steps for different classes of injection attacks.
Links between steps are labeled with the different attacks that use that path,
while the numbers next to attack labels denote the 
sequence of steps for a particular attack.
Most steps are common in all attacks. This
is because they are based on similar attack vectors as we mentioned in
the previous section. Large `X' marks on transitions 
indicate the points where defenses detect or prevent attacks.

@@

In this section we enumerate key observations
based on the dimensions we set in Section~\ref{sec:dimensions}.

@@

In this section we provide some lessons learned from
our observations and provide some recommendations
that developers of new mechanisms may find helpful.

>>>>

There is already a survey on mitigating software vulnerabilities in
general~\cite{SZ12}. The scope of that research is much broader that
ours and leaves out the majority of the approaches and mechanisms that
we consider in this work. This is mainly because that work includes
countermeasures that identify vulnerabilities using program analysis
(which takes place during the development or testing phases), while we
consider protection mechanisms that counter web attacks when they take
place. Similarly, static or dynamic analysis techniques that examine
applications to identify vulnerabilities that may lead to the attacks
that we described earlier are out of scope as well. While
countermeasures against {\sc sql} injection attacks have been
previously surveyed in a 2006 paper~\cite{HVO06}, a significant number
of new {\sc sql} injection defenses has been proposed since then, as
we mentioned earlier.

@@

Lack of input validation is a major vulnerability behind dangerouw web
application attacks. By taking advantage of this, attackers can inject
their code into applications to perform malicious tasks. Exploits of
this kind can have different forms depending on the execution context
of the application and the location of the programming flaw that leads
to the attack.

@@

Code injection attacks can be divided in two categories. The first
involves binary code and the second source language code. An
extensive survey on binary code injection attacks was conducted by
Lhee and Chapin~\cite{LC03}. Advances in memory corruption vulnerability
exploitation have been studied extensively~\cite{PB04, SPWS13}
and countermeasures to such attacks
have already been surveyed~\cite{YJP12}\cite[Sec. 13.8]{DKZ12}.
In this work
we do not consider binary code injection, focusing instead on defenses
that protect web applications against attacks based on the
injection of source code.

Figure~\ref{fig:taxonomy} presents a taxonomy of source
code injection attacks against web applications.
Such attacks may involve high-level language code, written in either a
{\it Domain Specific Language} ({\sc dsl}) or a {\it Dynamic Language}.
To illustrate, we briefly discuss examples from both categories.

Injection attacks that involve {\sc dsl}s constitute an important
subset of the source code injection problem, as {\sc dsl}s such as
{\sc sql} and {\sc xml} play a significant role in the development of
both web and mobile applications. For example, many applications have
interfaces through which a user enters input to interact with the
application, thereby interacting with the underlying database. This
input can become part of an {\sc sql} query and get executed on the
target database. Code injection attacks that exploit vulnerabilities
in database interfaces by taking advantage of input validation issues,
such as incorrectly passed parameters or incorrect type handling, are
called {\sc sql} injection attacks~\cite{CERT02,MS09,HVO06,SW06}.
Consider a trivial exploit that takes advantage of incorrectly
filtered quotation characters in an application that sends a temporary
password for a forgetful user via email by executing the following
query:

@@

\noindent
Attackers that would input the string {\tt anything' OR 'x'='x}
thereby obtaining the administrator's password by email. Savvy
programmers can use certain {\sc api} functions, such as {\sc php}'s
{\tt mysql\_real\_escape\_string()}, to detect malformed input, or,
better, use prepared {\sc sql} statements instead of statement
templates. Unfortunately, the increasing number of {\sc sql} injection
attacks suggests that programmers are not always that careful. Using
similar techniques, malicious users can mount other exploits based on
{\sc dsl}s such as {\sc xp}ath~\cite{SW06,CDL07,MKS09}, {\sc
  xml}~\cite{MSM13} and {\sc json}~\cite{SMS13}. The effects can be
wide-ranging. A malicious user can view sensitive information, destroy
or modify protected data, or even crash the entire application.

{\sc html} is another {\sc dsl} that can be used for malicious
purposes when an application does not properly handle user-supplied
data. Based on this vulnerability attackers can supply valid {\sc
  html}, typically via a parameter value, and inject their own content
into the page. {\sc html} injection is mainly associated with {\sc
  xss} attacks~\cite{BJM08,LZRL09}. However, it can also be used as a
vehicle for cross-site request forgery ({\sc csrf}) attacks in the
following manner: Consider a vulnerable bulletin board system where
{\tt img} tags are not checked. A malicious user could embed a {\sc
  csrf} request within an {\tt img} tag in the following manner:

@@

JavaScript injection attacks make up a large subset of dynamic
language code injection attacks. Such attacks are enabled when
a web application accepts
and redisplays data of uncertain origin without
proper validation and filtering. Based on this flaw, an attacker
can manage to inject a script in the JavaScript engine of a browser
and alter its execution flow~\cite{ELX07}.
JavaScript injection attacks are considered a critical issue
in web application security mainly because it is associated with
major vulnerabilities such as {\sc xss} attacks~\cite{SG07} and
Cross-Channel Scripting ({\sc xcs}) attacks~\cite{W10,BBB09}.
For a typical {\sc xss} example that involves JavaScript injection
consider a web page that prints the value
of a query parameter ({\tt query}) from the
page's {\sc url} as part of the page's content
without escaping the value. Attackers
can take advantage of this and inject an {\tt iframe} tag
into the page to steal a user's cookie and
send it via an image request to a web site
under their control (malicious.com).
This could be achieved by including the following
link to the malicious web site (or sending it via phishing
email) and inducing the user to click on it:

@@

We provide a step-by-step exploitation model to understand the process
of carrying out source code injection attacks against web
applications. Figure~\ref{fig:attacks} presents the required steps for
different classes of injection attacks. Links between steps are
labeled with the different attacks that use that path, while the
numbers next to attack labels denote the sequence of steps for a
particular attack. Most steps are common in all attacks. This is
because they are based on similar attack vectors as we mentioned in
the previous section. Large `X' marks on transitions indicate the
points where defenses detect or prevent attacks.

@@

We group our key observations on the~34 publications along the
dimensions we identified in Section~\ref{sec:dimensions}.

@@

Our observations lead to some lessons and recommendations that
developers of new mechanisms may find helpful. In particular, they
call for improving accuracy tests and code availability, while
aiming to reduce computational and deployment overheads.
